### 2023-10-16 notes/notes.tex
Astropy Table
I found that the astropy package supports the mrt format and can make my life easier

% Created 2023-10-11 Wed 19:37
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\date{\today}
\title{Karachentsev List notes}
\hypersetup{
 pdfauthor={},
 pdftitle={Karachentsev List notes},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 28.3 (Org mode 9.6.1)}, 
 pdflang={English}}
\usepackage{biblatex}
\addbibresource{/home/dp/Documents/Research_paper_SFR/My Library/My Library.bib}
\begin{document}

\maketitle
\tableofcontents

This list contains measurements from the galaxies in the Local Cosmological Volume (LCV), and the last update of data is on 2023-06-30. I downloaded the database from \autocite{karachentsevUPDATEDNEARBYGALAXY2013}

\section{Original data set}
\label{sec:orgbd60bab}

The data I will use for this project come from the Karachentsev  catalog (2022-12-02), which is an updated version of the list used in Kroupa et al. 2020
From the list I use the tables ``Catalog of Nearby Galaxies'', ``Global Parameters of the Nearby Galaxies'' and ``List of the nearby galaxies with measured SFR''

\subsubsection{{\bfseries\sffamily IDEA} I want to merge the coordinate columns into one}
\label{sec:org9296394}

\subsection{Join the tables}
\label{sec:org3cf69e8}

<Table length=1436>
  name    dtype       unit                       description                      class     n\textsubscript{bad}
-------- ------- ------------- ----------------------------------------------- ------------ -----
    Name   str18                            Galaxy name in well-known catalogs       Column     0
   RAh\textsubscript{1}   int64             h                 Hour of Right Ascension (J2000)       Column     0
   RAm\textsubscript{1}   int64           min               Minute of Right Ascension (J2000)       Column     0
   RAs\textsubscript{1} float64             s               Second of Right Ascension (J2000)       Column     0
   DE-\textsubscript{1}    str1                               Sign of the Declination (J2000)       Column     0
   DEd\textsubscript{1}   int64           deg                   Degree of Declination (J2000)       Column     0
   DEm\textsubscript{1}   int64        arcmin                Arcminute of Declination (J2000)       Column     0
   DEs\textsubscript{1}   int64        arcsec                Arcsecond of Declination (J2000)       Column     0
   a26\textsubscript{1} float64        arcmin                      Major angular diameter (1) MaskedColumn    11
     b/a float64                                      Apparent axial ratio (1) MaskedColumn    11
    AB\textsubscript{1} float64           mag               Galactic extinction in B band (2) MaskedColumn     1
l\textsubscript{FUVmag}    str1                                          Limit flag on FUVmag MaskedColumn  1104
  FUVmag float64           mag                    GALEX FUV band magnitude (3) MaskedColumn   311
    Bmag float64           mag                   Integral B band magnitude (4) MaskedColumn     8
 l\textsubscript{Hamag}    str1                                           Limit flag on Hamag MaskedColumn  1306
   Hamag float64           mag   Integral H\{alpha\} line emission magnitude (5) MaskedColumn   700
    Kmag float64           mag                   2MASS K\textsubscript{S}\_ band magnitude (6) MaskedColumn    11
  f\textsubscript{Kmag}    str1                                          [*] Flag on Kmag (7) MaskedColumn   362
 l\textsubscript{21mag}    str1                                           Limit flag on 21mag MaskedColumn  1226
   21mag float64           mag                    H I 21 cm line magnitude (8) MaskedColumn   494
     W50   int64        km / s     H I line with at 50\% level from maximum (9) MaskedColumn   623
   TType   int64                                     Morphology type code (10) MaskedColumn     4
    Tdw1    str5                                  Dwarf galaxy morphology (11) MaskedColumn   233
    Tdw2    str1               Dwarf galaxy surface brightness morphology (12) MaskedColumn   250
    RVel   int64        km / s               Heliocentric radial velocity (13) MaskedColumn   454
     Dis float64           Mpc                                        Distance       Column     0
   f\textsubscript{Dis}    str4                           Method flag used to obtain Dis (14)       Column     0
   RAh\textsubscript{2}   int64             h                 Hour of Right Ascension (J2000)       Column     0
   RAm\textsubscript{2}   int64           min               Minute of Right Ascension (J2000)       Column     0
   RAs\textsubscript{2} float64             s               Second of Right Ascension (J2000)       Column     0
   DE-\textsubscript{2}    str1                               Sign of the Declination (J2000)       Column     0
   DEd\textsubscript{2}   int64           deg                   Degree of Declination (J2000)       Column     0
   DEm\textsubscript{2}   int64        arcmin                Arcminute of Declination (J2000)       Column     0
   DEs\textsubscript{2}   int64        arcsec                Arcsecond of Declination (J2000)       Column     0
   a26\textsubscript{2} float64           kpc                       Major linear diameter (1) MaskedColumn    11
     inc   int64           deg                                     Inclination MaskedColumn    11
      Vm   int64        km / s            Amplitude of rotational velocity (2) MaskedColumn   635
    AB\textsubscript{2} float64           mag                  Internal B band extinction (3) MaskedColumn    11
    BMag float64           mag                   Absolute B band magnitude (4) MaskedColumn     8
     SBB float64 mag / arcsec2           Average B band surface brightness (5) MaskedColumn    11
 logKLum float64     dex(Lsun)                    Log K\textsubscript{S}\_ band luminosity (6) MaskedColumn    10
  logM26 float64     dex(Msun)             Log mass within Holmberg radius (7) MaskedColumn   634
l\textsubscript{logMHI}    str1                                          Limit flag on logMHI MaskedColumn  1226
  logMHI float64     dex(Msun)                           Log hydrogen mass (8) MaskedColumn   494
     VLG   int64        km / s                             Radial velocity (9) MaskedColumn   454
  Theta1 float64                                              Tidal index (10) MaskedColumn    76
      MD   str19                                      Main disturber name (11) MaskedColumn    76
  Theta5 float64                                      Another tidal index (12) MaskedColumn    76
  Thetaj float64      dex(---)              Log K band luminosity density (13) MaskedColumn   180
\end{document}