### 2024-09-30 compare/compare.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
  \setmainfont[]{Latin Modern Roman}
  \setmathfont[]{Latin Modern Math}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{luacolor}
  \usepackage[soul]{lua-ul}
\else
  \usepackage{soul}

\fi
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Comparison of Catalogs},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\newcommand{\runninghead}{A Preprint }
\title{Comparison of Catalogs}
\def\asep{\\\\\\ } % default: all authors on same column
\author{}
\date{}
\begin{document}
\maketitle

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}
\section{The data}\label{the-data}

In this script we will compare 2 catalogs Kovlakas et al. (2021) and
Karachentsev and Kaisina (2013)

\begin{itemize}
\tightlist
\item
  The data have been joined based on their position in the sky (Ra,
  Dec).

  \begin{itemize}
  \tightlist
  \item
    We assume that every galaxy within 2 arc seconds of the initial
    coordinates is the same galaxy.
  \end{itemize}
\item
  We use TOPCAT to create two joins, an inner and an outer join
\item
  We will use the inner join for 1-1 comparisons
\item
  If we see that the data are similar we can use the outer join
\item
  For the comparison we keep the parameters names exactly they are given
  in the catalogs
\end{itemize}

The dataset we are going to use for the comparison (inner join) consists
of 288 galaxies and 168 columns.

\section{Catalog Completeness}\label{catalog-completeness}

Checking for completeness in galaxy catalogs is essential to ensure that
the data accurately represents the true population of galaxies.
Incomplete catalogs can lead to biased results in statistical studies,
such as the distribution of galaxy luminosity, mass, or star formation
rates. Additionally, missing galaxies, especially those at faint
magnitudes or large distances, can distort cosmological measurements and
hinder our understanding of galaxy formation and evolution.

Completeness checks are crucial for addressing selection biases,
identify gaps in the data and guide follow-up observations, ensuring
that the catalog provides a reliable sample for scientific analysis.
Without these checks, conclusions drawn from the data may be inaccurate
or incomplete.

\subsection{Completeness of the LCV
Catalog}\label{completeness-of-the-lcv-catalog}

The local volume selection has been made by taking into account galaxies
with: 1. Radial velocities of
\begin{equation}\phantomsection\label{eq-lcv-vel}{
    V_{LG}<600 \text{ km}\cdot\text{m}^{-1}
    }\end{equation} 2. Distances of D\textless11 Mpc

A simultaneous fulfillment of both conditions (1) and (2) is not
required.

\begin{itemize}
\tightlist
\item
  Completeness within a 10 Mpc radius is difficult to assess due to:

  \begin{itemize}
  \tightlist
  \item
    Variability in galaxy properties (luminosity, size, surface
    brightness, gas content)
  \item
    Errors in distance measurements (Tully--Fisher method errors of
    \textasciitilde20--25\%), especially at the 10 Mpc boundary.

    \begin{itemize}
    \tightlist
    \item
      Accurate distances are mainly known within \textasciitilde5 Mpc.
    \end{itemize}
  \item
    Non-Hubble motions (\textasciitilde300 km/s) may make up half of the
    adopted velocity constraint Equation~\ref{eq-lcv-vel}

    \begin{itemize}
    \tightlist
    \item
      Solution for our usage: only keep the galaxies inside a radius of
      \(D=11 Mpc\)
    \item
      HI sources in surveys with low angular resolution: ``The presence
      around our Galaxy of hundreds of high-velocity clouds with low
      line-of-sight velocities and small W50 widths also provokes the
      inclusion of false ``nearby'' dwarf galaxies in the
      LV''Karachentsev, Makarov, and Kaisina (2013)
    \item
      photographic emulsion defect -\textgreater{} exotic cases of
      galaxies -\textgreater{} radial velocity of \(+614 km s^{−1}\)
    \end{itemize}
  \end{itemize}
\item
  Astro-Spam and Survey Errors:

  \begin{itemize}
  \tightlist
  \item
    Automatic surveys produce false detections (e.g., stars
    misclassified as galaxies, high-velocity clouds mistaken for
    dwarfs).
  \end{itemize}
\item
  Conditional Completeness Estimate:

  \begin{itemize}
  \tightlist
  \item
    Galaxies brighter than \(M_B^c = -11^m\) or with linear diameters
    larger than \(A_{26} = 1.0\) kpc show (40--60)\% completeness.
  \item
    ``among the members of the Local Group (D \textless{} 1 Mpc), only
    half of the galaxies have absolute magnitudes brighter than −11m.
    Consequently, more than half of the ultra-faint dwarf companions
    around normal galaxies, like the Sombrero galaxy (D = 9.3 Mpc),
    still remain outside our field of view.''
  \end{itemize}
\item
  Undetected Ultra-Faint Dwarfs:

  \begin{itemize}
  \tightlist
  \item
    Many faint dwarf galaxies remain undetected beyond the Local Group.
    Estimated population of undetected dwarfs could be as large as
    \(10^3–10^4\) within the LV.
  \end{itemize}
\item
  Surface Brightness Distribution:

  \begin{itemize}
  \tightlist
  \item
    Surface brightness remains consistent across distances, except for
    ultra-faint dwarfs (\(SB <31 \text{mag}\cdot\text{arcsec}^{-2}\)).
  \item
    Faintest dwarf galaxies are detectable only nearby due to their
    resolution into individual stars.
  \end{itemize}
\item
  Luminosity-Size Relationship:

  \begin{itemize}
  \tightlist
  \item
    Observations align with cosmological models predicting \(L\sim A^3\)
    Navarro, Frenk, and White (1996), though deviations occur for
    extremely low surface brightness galaxies.
  \item
    ``The deviation from it at the extremely low surface brightness end
    is due to a systematic overestimation of dwarf galaxy sizes, the
    brightness profiles of which lie entirely below the Holmberg
    isophote.''
  \end{itemize}
\end{itemize}

\subsection{Completeness of HECATE}\label{completeness-of-hecate}

The completeness of HECATE is difficult to assess due to: - Unknown
selection function of HyperLEDA and selection effects from other
cross-correlated catalogs. - Estimation based on comparing B-band
luminosity distribution with the galaxy luminosity function (LF).

\begin{itemize}
\item
  HECATE is:

  \begin{itemize}
  \item
    Complete down to \(L_B\sim 10^{9.5} L_{B,\odot}\) for distances less
    than 33 Mpc.
  \item
    Complete down to \(L_B\sim 10^{10} L_{B,\odot}\) for distances
    between 67 Mpc and 100 Mpc.
  \item
    Incomplete at distances greater than 167 Mpc, even for the brightest
    galaxies.
  \end{itemize}
\item
  Completeness estimates based on B-band luminosity density:

  \begin{itemize}
  \item
    \textgreater75\% complete for distances less than 100 Mpc.
    \(\sim 50\%\) complete at distances of \(\sim170\) Mpc.
  \item
    Completeness exceeds 100\% within 30 Mpc due to the overdensity of
    galaxies around the Milky Way.
  \end{itemize}
\item
  Completeness in terms of stellar mass (M*):

  \begin{itemize}
  \tightlist
  \item
    Similar to B-band completeness when measured with K\_s-band
    luminosity as a tracer for stellar mass. Overdensity at small
    distances and cut-off at large distances are observed.
  \end{itemize}
\item
  Completeness in terms of star formation rate (SFR):

  \begin{itemize}
  \item
    \textasciitilde50\% complete between 30 and 150 Mpc.
  \item
    Lower SFR completeness due to limitations in WISE-based SFR
    estimates, which lack full sky coverage.
  \item
    Despite IRAS's limited depth, it provides \textgreater50\% coverage
    for star-forming galaxies in the local neighborhood.
  \end{itemize}
\end{itemize}

In this section we will check the completeness

\begin{longtable}[]{@{}lr@{}}
\toprule\noalign{}
Table & Number of galaxies \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Inner join & 288 \\
Outer join & 2901 \\
LCV & 1316 \\
HECATE & 2901 \\
Unique galaxies in LCV & 1028 \\
Unique Galaxies in Hecate & 2613 \\
\end{longtable}

\subsection{Completeness of the Inner
join}\label{completeness-of-the-inner-join}

\[
\text{Completeness (X)}=\frac{\text{(Galaxies in Inner Join)}}{\text{(Galaxies in X)}}×100\%
\]

Completeness (HECATE)= 10 \%

Completeness (LCV)= 22 \%

\subsection{Completeness in Outer
join}\label{completeness-in-outer-join}

\[
\text{Completeness (X)}=\frac{\text{(Galaxies in Outer Join form X)}}{\text{(Galaxies in X)}}×100\%
\]

Completeness (HECATE)= 90 \%

Completeness (LCV)= 78 \%

Combined Completeness
=\(\frac{\text{Total galaxies in Outer}}{\text{Unique galaxies in HECATE + LCV}}\)=
80 \%

\subsection{Completeness of the Data}\label{completeness-of-the-data}

\subsubsection{Distance}\label{distance}

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{compare_files/figure-pdf/fig-dis-comp-output-1.pdf}

}

\subcaption{\label{fig-dis-comp-1}HECATE}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{compare_files/figure-pdf/fig-dis-comp-output-2.pdf}

}

\subcaption{\label{fig-dis-comp-2}LCV}

\end{minipage}%

\caption{\label{fig-dis-comp}Histograms showing the Distance
Completeness of the Catalogs}

\end{figure}%

\begin{figure}

\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{compare_files/figure-pdf/fig-type-comp-output-1.pdf}

}

\subcaption{\label{fig-type-comp-1}HECATE}

\end{minipage}%
%
\begin{minipage}{0.50\linewidth}

\centering{

\includegraphics{compare_files/figure-pdf/fig-type-comp-output-2.pdf}

}

\subcaption{\label{fig-type-comp-2}LCV}

\end{minipage}%

\caption{\label{fig-type-comp}Histograms showing the Type Completeness
of the Catalogs}

\end{figure}%

As we can see from the histograms Figure~\ref{fig-dis-comp} and
Figure~\ref{fig-type-comp} the sample of unique galaxies of each
catalog, gets smaller by an almost constant proportion (Inner join).

This means there is no bias in the selection of the galaxies.

\section{How are we going to compare the
data?}\label{how-are-we-going-to-compare-the-data}

\subsection{\texorpdfstring{Scatter plots and \(R^2\)
calculation}{Scatter plots and R\^{}2 calculation}}\label{scatter-plots-and-r2-calculation}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \(R^2\): Measures the proportion of variance explained by the linear
  model.
\item
  Slope of the Fitted Line: Should be close to 1 for a 1-1
  correlation.\footnote{Some data seem to have a very good linear
    correlation but they have many outliers. This is why we will clip
    the outliers with \(\sigma > 3\)}
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  Pearson Correlation \(\rho\): Measures the strength and direction of
  the linear relationship between two variables, ranging from -1 to 1.
  \footnote{In simple linear regression, \(R^2\) is the square of the
    Pearson correlation coefficient \(\rho\).}
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  \ul{Plots}: Plots are essential for visually assessing the
  relationship between two datasets, identifying correlations, trends,
  and outliers, and evaluating the fit of linear models.
\end{enumerate}

\begin{itemize}
\tightlist
\item
  Histograms: Because not all of our data have the same number of
  counts, the comparison with histograms between data that are not the
  same, doesn't help us right now.\footnote{When we will use the outer
    join table we could use histograms due to the large number of
    counts.} This is why we will only use histograms for comparing the
  distribution of same-data columns normalized by their maximum value
\end{itemize}

\begin{itemize}
\item
  Correlation Heatmaps: A correlation heatmap is a graphical tool that
  displays the correlation between multiple variables as a color-coded
  matrix. It's like a color chart that shows us how closely related
  different variables are. In a correlation heatmap, each variable is
  represented by a row and a column, and the cells show the correlation
  between them. The color of each cell represents the strength and
  direction of the correlation, with darker colors indicating stronger
  correlations.
\item
  Kernel Density Estimate (KDE) plot: The KDE plot visually represents
  the distribution of data, providing insights into its shape, central
  tendency, and spread.
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  Percentage change: We can calculate the percentage change of the data
  for each galaxy and then we can see if the data are similar, based on
  minimum, the maximum and the mean value of the difference.
\end{enumerate}

\[\text{Percentage change} = \frac{V_{Hecate} - V_{LCV}}{V_{Hecate}}\cdot 100 \%\]

\section{Comparable data}\label{comparable-data}

\subsection{Coordinates}\label{coordinates}

\begin{longtable}[]{@{}cccc@{}}
\toprule\noalign{}
LCV & HECATE & Description & Pearson Correlation {[}-1,1{]} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Ra\_1 & RA\_2 & Right Ascension & 1.0 \\
Dec\_1 & DEC\_2 & Declination & 1.0 \\
Dis & D & Distance & 0.881 \\
\end{longtable}

\subsubsection{Right Ascension}

\includegraphics{compare_files/figure-pdf/cell-13-output-1.pdf}

\subsubsection{Declination}

\includegraphics{compare_files/figure-pdf/cell-14-output-1.pdf}

\subsubsection{Distance}

\includegraphics{compare_files/figure-pdf/cell-15-output-1.pdf}

\subsection{Velocities}\label{velocities}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2466}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2603}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2466}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2466}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
LCV
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
HECATE
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Description
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Linear Correlation
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
RVel (km/s) & V (km/s) & Heliocentric radial velocity & 0.994 \\
VLG (km/s) & & Radial velocity & \\
cz (km/s) & & Heliocentric velocity & \\
& V\_VIR (km/s) & Virgo-infall corrected radial velocity & \\
\end{longtable}

\includegraphics{compare_files/figure-pdf/cell-17-output-1.pdf}

\includegraphics{compare_files/figure-pdf/cell-19-output-1.pdf}

{[}?{]} The close correlation between all of the velocities, could be
due to the fact that all of them measure the velocity of each galaxy,
but from a different frame of reference.

\subsection{Morphology and Geometry}\label{morphology-and-geometry}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2466}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2466}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2603}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2466}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
LCV
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
HECATE
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Description
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pearson Correlation {[}-1,1{]}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
TType & T (with errors) & Numerical Hubble type following the de
Vaucouleurs system & 0.7107 \\
inc & INCL & Inclination (deg) & 0 \\
a26\_1 (Major) & R1 (Semi-major axis) & angular diameter (arcmin) & 0 \\
\end{longtable}

\subsubsection{Galaxy Types}

\includegraphics{compare_files/figure-pdf/cell-22-output-1.pdf}

``Morphological type of galaxy in the numerical code according to the
classification by de Vaucouleurs et al.~(1991). It should be noted that
about three quarters of objects in the LV are dwarf galaxies, which
require a more detailed morphological classification. For example, dwarf
spheroidal galaxies and normal ellipticals are usually denoted by the
same numerical code T \textless{} 0, although their physical properties
drastically differ. The classification problem arises as well for the
``transient'' type dwarf galaxies, T r, which combine the features of
spheroidal (Sph) and irregular (Ir)systems. Due to small classification
errors, such objects may ``jump'' from one end of the T scale to the
other.'' Karachentsev, Makarov, and Kaisina (2013)

\begin{itemize}
\tightlist
\item
  we can assume that the galaxies from the upper left and lower right
  regions of the plot are classified differently because of this
\item
  this might explain the large errors E\_T of HECATE
\item
  why we see a difference in the distribution of
  Figure~\ref{fig-type-comp}
\end{itemize}

We can remove the ``problematic'' galaxies by only keeping the ones
with: \[
|{T_{HECATE}-T_{LCV}}|<7
\]

\includegraphics{compare_files/figure-pdf/cell-23-output-1.pdf}

\subsubsection{Inclination}

\begin{figure}

\begin{minipage}{0.50\linewidth}
\includegraphics{compare_files/figure-pdf/cell-29-output-1.pdf}\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\includegraphics{compare_files/figure-pdf/cell-29-output-2.pdf}\end{minipage}%
\newline
\begin{minipage}{\linewidth}

\begin{longtable}[]{@{}llll@{}}
\toprule\noalign{}
& inc & INCL & Percentage Change {[}\%{]} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
count & 287 & 209 & 202 \\
mean & 60 & 59 & -0 \\
std & 19 & 23 & 0 \\
min & 9 & 0 & -3 \\
25\% & 47 & 47 & -0 \\
50\% & 60 & 59 & 0 \\
75\% & 72 & 78 & 0 \\
max & 90 & 90 & 1 \\
\end{longtable}

\end{minipage}%

\end{figure}%

We can see that for values in the range
\([\sim 30^\circ,\sim 80^\circ]\), the values of the LCV inclination are
higher. However, since their means, median, min and maxes are similar
and the percentage change is practically 0\% (mean, median, \(\sigma\) =
0 with a range \([-3\%,1\%]\)), we can ignore the differences and assume
they are the same values.

\subsubsection{Major Axis}

\includegraphics{compare_files/figure-pdf/cell-30-output-1.pdf}

it is not very clear if we truly have a correlation or not. We need to
see the linear correlation of the decimal logarithms.

\includegraphics{compare_files/figure-pdf/cell-32-output-1.pdf}

\subsection{Luminosities}\label{luminosities}

\begin{longtable}[]{@{}cccc@{}}
\toprule\noalign{}
LCV & HECATE & Description & Pearson Correlation {[}-1,1{]} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
logKLum & logL\_K & & 0 \\
\end{longtable}

\includegraphics{compare_files/figure-pdf/cell-34-output-1.pdf}

\begin{longtable}[]{@{}llll@{}}
\toprule\noalign{}
& log(L\_K)\_\{LCV\}\$ & log(L\_K)\_\{HEC\} & Percentage Change
{[}\%{]} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
count & 287 & 70 & 70 \\
mean & 8 & 9 & -4 \\
std & 1 & 1 & 8 \\
min & 3 & 5 & -48 \\
50\% & 8 & 9 & -0 \\
max & 11 & 11 & 7 \\
\end{longtable}

\subsection{Magnitudes}\label{magnitudes}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2466}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2466}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2466}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2603}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
LCV
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
HECATE
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Description
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pearson Correlation {[}-1,1{]}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
mag\_B (with errors) & BT (with errors) & & 0 \\
Kmag & K & 2MASS band magnitude (both) & 0 \\
\end{longtable}

\subsubsection{B mag}

\includegraphics{compare_files/figure-pdf/cell-38-output-1.pdf}

\subsubsection{K mag}

\includegraphics{compare_files/figure-pdf/cell-39-output-1.pdf}

\begin{longtable}[]{@{}lll@{}}
\toprule\noalign{}
& Percentage Change {[}\%{]} & Percentage Change, after 3 sigma clip
{[}\%{]} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
count & 70 & 70 \\
mean & -5 & -5 \\
std & 12 & 10 \\
min & -70 & -42 \\
50\% & -0 & -0 \\
max & 16 & 16 \\
\end{longtable}

\includegraphics{compare_files/figure-pdf/cell-41-output-1.pdf}

{[}?{]}

\subsection{SFR}\label{sfr}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2466}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2466}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2603}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2466}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
LCV
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
HECATE
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Description
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Count
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
& logSFR\_TIR & Decimal logarithm of the total-infrared SFR estimate
{[}Msol/yr{]} & 21 \\
& logSFR\_FIR & Decimal logarithm of the far-infrared SFR estimate
{[}Msol/yr{]} & 22 \\
& logSFR\_60u & Decimal logarithm of the 60um SFR estimate {[}Msol/yr{]}
& 48 \\
& logSFR\_12u & Decimal logarithm of the 12um SFR estimate {[}Msol/yr{]}
& 26 \\
& logSFR\_22u & Decimal logarithm of the 22um SFR estimate {[}Msol/yr{]}
& 23 \\
& logSFR\_HEC & Decimal logarithm of the homogenised SFR estimate
{[}Msol/yr{]} & 73 \\
& logSFR\_GSW & Decimal logarithm of the SFR in GSWLC-2 {[}Msol/yr{]} &
0 \\
SFRFUV & & FUV derived integral star formation rate & 220 \\
SFRHa & & H\{alpha\} derived integral star formation rate & 223 \\
\end{longtable}

\includegraphics{compare_files/figure-pdf/cell-43-output-1.pdf}

\subsection{Masses}\label{masses}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
LCV
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
HECATE
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Description
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Count
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
logM26 & & Log mass within Holmberg radius & 233 \\
logMHI & & Log mass within Holmberg radius & 233 \\
& logM\_HEC & Decimal logarithm of the stellar mass {[}Msol{]} & 64 \\
& logM\_GSW & Decimal logarithm of the stellar mass in GSWLC-2
{[}Msol{]} & 0 \\
logStellarMass & & Stellar Mass from \(M_*/L=0.6\) & 287 \\
\end{longtable}

\subsubsection{Stellar Masses Comparison}

\includegraphics{compare_files/figure-pdf/cell-45-output-1.pdf}

\begin{longtable}[]{@{}ll@{}}
\toprule\noalign{}
& Percantage Change {[}\%{]} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
count & 64 \\
mean & -1 \\
std & 5 \\
min & -19 \\
50\% & 1 \\
max & 9 \\
\end{longtable}

\subsubsection{Heatmap}

\includegraphics{compare_files/figure-pdf/cell-48-output-1.pdf}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-karachentsevSTARFORMATIONPROPERTIES2013a}
Karachentsev, Igor D., and Elena I. Kaisina. 2013. {``{STAR FORMATION
PROPERTIES IN THE LOCAL VOLUME GALAXIES VIA Hα AND FAR-ULTRAVIOLET
FLUXES}.''} \emph{The Astronomical Journal} 146 (3): 46.
\url{https://doi.org/10.1088/0004-6256/146/3/46}.

\bibitem[\citeproctext]{ref-karachentsevUPDATEDNEARBYGALAXY2013}
Karachentsev, Igor D., Dmitry I. Makarov, and Elena I. Kaisina. 2013.
{``{UPDATED NEARBY GALAXY CATALOG}.''} \emph{The Astronomical Journal}
145 (4): 101. \url{https://doi.org/10.1088/0004-6256/145/4/101}.

\bibitem[\citeproctext]{ref-kovlakasHeraklionExtragalacticCatalogue2021}
Kovlakas, K., A. Zezas, J. J. Andrews, A. Basu-Zych, T. Fragos, A.
Hornschemeier, K. Kouroumpatzakis, B. Lehmer, and A. Ptak. 2021. {``The
{Heraklion Extragalactic Catalogue} ({HECATE}): A Value-Added Galaxy
Catalogue for Multimessenger Astrophysics.''} \emph{Monthly Notices of
the Royal Astronomical Society} 506 (September): 1896--1915.
\url{https://doi.org/10.1093/mnras/stab1799}.

\bibitem[\citeproctext]{ref-navarroStructureColdDark1996}
Navarro, Julio F., Carlos S. Frenk, and Simon D. M. White. 1996. {``The
{Structure} of {Cold Dark Matter Halos}.''} \emph{The Astrophysical
Journal} 462 (May): 563. \url{https://doi.org/10.1086/177173}.

\end{CSLReferences}

\end{document}