### 2025-03-16 template/thesis_01_titlepage.tex

\begin{titlepage}
%
%
\begin{figure*}[ht]
\centering
\includegraphics[width=0.95\textwidth]{logo_header.png}
\end{figure*}

\vspace{10mm}
\hbox{ % Horizontal box
\hspace*{0.08\textwidth} % Whitespace to the left of the title page
\rule{1.5pt}{0.80\textheight} % Vertical line
\hspace*{0.02\textwidth} % Whitespace between the vertical line and title page text
\parbox[b]{0.90\textwidth}{ % Paragraph box which restricts text to less than the width of the page

{\large \textsc{Πτυχιακή Εργασία}}\\[\baselineskip] % Tagline or further description
{\noindent\Large\bfseries
title
}\\[2\baselineskip] % Title

{\Large Doctor Who}\\ % AUTHOR's name
{Α/Μ~201800142}\\[3\baselineskip] % Registrar number for the AUTHOR

{\em Επιβλέπων:}\\[0.25\baselineskip]
{\large Despina Hatzidimitriou}\\[0.25\baselineskip] % SUPERVISOR's name
{\large Professor} % SUPERVISOR's Title

\vspace{0.15\textheight} % Whitespace between the blocks
{\noindent Athens 2025}\\[\baselineskip] % Place / Year
}}
%
%
\end{titlepage}

\newpage
\thispagestyle{empty}
\phantom{~}
\vfill